Bibliography

### 2016-11-24 Draft-1.tex
Conditional comparison, Recursor function

The elusive predecessor function was created by Stephen Kleene, a student of Church's, and was something of a landmark in the development of $\lambda$-calculus.
Firstly, the function transforms the pair (n, n-1) (the argument $p$) into the pair (n+1, n) (the new pair's elements are an increment and a copy of the first value respectively):
\begin{equation*}
\bm{\psi} = \lambda pc.c(S(pT))(pT)
\end{equation*}
To find the predecessor to $n$ therefore, we apply this function $n$ times to the pair (0, 0) and the pair (n, n-1) is formed, after which the second element can easily be selected.
\begin{equation*}
\bm{\mathrm{P}} = \lambda n.n \bm{\psi} (\lambda x.00) F 
\end{equation*}
\begin{equation*}
\begin{aligned}
\bm{\mathrm{P3}} &= (\lambda n.n [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}] \bm{\mathrm{F}}) [\lambda sz.s(s(s(z)))]\\
&= ([\lambda sz.s(s(s(z)))] [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}]) \bm{\mathrm{F}}\\
&= ([\lambda sz.s(s(s(z)))] [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}]) \bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})][\lambda x.\bm{\mathrm{00}}]))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})](\lambda c.c(\bm{\mathrm{S}}[\lambda x.\bm{\mathrm{00}}]\bm{\mathrm{T}})([\lambda x.\bm{\mathrm{00}}]\bm{\mathrm{T}})))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})](\lambda c.c\bm{\mathrm{10}}))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})][\lambda c.c\bm{\mathrm{21}}]\bm{\mathrm{F}}\\
&= [\lambda c.c\bm{\mathrm{32}} ])\bm{\mathrm{F}}\\
&= \bm{\mathrm{2}}
\end{aligned}
\end{equation*}
Of course now the subtraction function can be found (using \textbf{D} for difference):
\begin{equation*}
\bm{\mathrm{D}} = \lambda ab.bPa
\end{equation*}
Bearing in mind the predecessor of \textbf{0} is \textbf{0}, this will produce \textbf{0} if $b>a$.
\subsubsection{Further Conditional Logic}
With this useful quirk of the \textbf{D} function, we can produce comparative logical operators:
\begin{gather*}
\begin{aligned}
\bm{\leq} = \lambda ab.\bm{\mathrm{Z}}(Dab)
\end{aligned}
\end{gather*}
\end{document}

>>>>

The elusive predecessor function was created by Stephen Kleene, a student of Church's, and was something of a landmark in the development of $\lambda$-calculus.
Firstly, the function transforms the pair (n, n-1) (the argument $p$) into the pair (n+1, n) (the new pair's elements are an increment and a copy of the first value respectively):
\begin{equation*}
\bm{\psi} = \lambda pc.c(S(pT))(pT)
\end{equation*}
To find the predecessor to $n$ therefore, we apply this function $n$ times to the pair (0, 0) and the pair (n, n-1) is formed, after which the second element can easily be selected.
\begin{equation*}
\bm{\mathrm{P}} = \lambda n.n \bm{\psi} (\lambda x.00) F 
\end{equation*}
\begin{equation*}
\begin{aligned}
\bm{\mathrm{P3}} &= (\lambda n.n [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}] \bm{\mathrm{F}}) [\lambda sz.s(s(s(z)))]\\
&= ([\lambda sz.s(s(s(z)))] [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}]) \bm{\mathrm{F}}\\
&= ([\lambda sz.s(s(s(z)))] [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})] [\lambda x.\bm{\mathrm{00}}]) \bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})][\lambda x.\bm{\mathrm{00}}]))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})](\lambda c.c(\bm{\mathrm{S}}[\lambda x.\bm{\mathrm{00}}]\bm{\mathrm{T}})([\lambda x.\bm{\mathrm{00}}]\bm{\mathrm{T}})))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})]([\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})](\lambda c.c\bm{\mathrm{10}}))\bm{\mathrm{F}}\\
&= [\lambda pc.c(\bm{\mathrm{S}}(p\bm{\mathrm{T}}))(p\bm{\mathrm{T}})][\lambda c.c\bm{\mathrm{21}}]\bm{\mathrm{F}}\\
&= [\lambda c.c\bm{\mathrm{32}} ])\bm{\mathrm{F}}\\
&= \bm{\mathrm{2}}
\end{aligned}
\end{equation*}
Of course now the subtraction function can be found (using \textbf{D} for difference):
\begin{equation*}
\bm{\mathrm{D}} = \lambda ab.bPa
\end{equation*}
Bearing in mind the predecessor of \textbf{0} is \textbf{0}, this will produce \textbf{0} if $b>a$.
\subsubsection{Further Conditional Logic}

With this useful quirk of the \textbf{D} function, we can produce comparative logical operators:
\begin{equation*}
\begin{aligned}
\bm{\leq} &= \lambda ab.\bm{\mathrm{Z}}(\bm{\mathrm{D}}ab)
  & % seperation
\bm{\geq} &= \lambda ab.\bm{\mathrm{Z}}(\bm{\mathrm{D}}ba)
  \\
\bm{<} &= \lambda ab.\bm{\mathrm{Z}}(\bm{\mathrm{D}}a[Sb])
  &
\bm{>} &= \lambda ab.\bm{\mathrm{Z}}(\bm{\mathrm{D}}[Sb]a)
\end{aligned}
\end{equation*}
And through combination of these, we can create the equivalence and inequivalence operators (using $\bm{\doteq}$ for clarity from $=$:
\begin{equation*}
\begin{aligned}
\bm{\doteq} &= \lambda ab.\wedge(\bm{\leq}ab)(\bm{\geq}ab)\\
\bm{\neq} &= \lambda ab.\vee(\bm{<}ab)(\bm{>}ab)
\end{aligned}
\end{equation*}
\subsubsection{Recursion}

So far, Church's $\lambda$-Calculus is very close to the operation of a modern computation language. It has representation for boolean values, for conditional statements, for positive integers (signed integers can be expressed as a pair of a positive integer and a boolean for sign) and arithmetic, as well as set operations for booleans and comparators for integers - In fact, we are very close to our definition of computability save one thing. Continuity and repetition is one of the most significant aspects of computers, the power of this being the ability to transform a finite configuration into a potentially infinite result. This is done through one of the most remarkable functions in $\lambda$-Calculus, the recursor function:

\begin{equation*}
\bm{\mathrm{R}} = \lambda y.(\lambda x.y(xx))(\lambda x.y(xx)
\end{equation*}

\begin{equation*}
\begin{aligned}
\bm{\mathrm{R}}a &= (\lambda x.a(xx))[\lambda x.a(xx)]\\
&= a([\lambda x.a(xx)][\lambda x.a(xx)]\\
&= a(\bm{\mathrm{R}}a)
\end{aligned}
\end{equation*}
\end{document}
